Algorithmic Approaches to Match Degraded Land Impressions

### 2017-10-22 degraded-paper.Rmd
I don't remember what changes these were

custom <- list(
  theme_bw(),
  scale_linetype_manual(values=3:2),
  scale_colour_manual(values=c("black", "grey40")),
  ylab(expression(paste("Relative Height (in ",mu,m,")", sep=""))), 
  xlab(expression(paste("Relative Location (in ",mu,m,")", sep="")))
  )

p1 <- ggplot(data = final_profs, aes(x = y, y = l30, colour = profile_id)) +
    geom_line(aes(linetype=profile_id)) +
    custom

p2 <- ggplot(data = final_doublesmoothed, aes(x = y, y = l30, colour = profile_id)) +
    geom_line(aes(linetype=profile_id)) +
    geom_smooth(inherit.aes = FALSE, aes(x = y, y = avgl30), size = .5, se=FALSE, colour="black") +
    custom

p3 <- ggplot(data = final_doublesmoothed, aes(x = y, y = l50, colour = profile_id)) +
    geom_line(aes(linetype=profile_id)) +
    geom_hline(yintercept=0, size = .5, colour="black") +
    custom +
  ylab("Adjusted relative height")

@@

```{r roughcordist, echo=FALSE, out.width='.5\\textwidth',  fig.cap='Empirical CDFs of the Roughness Correlation compared to the CCF for known matches and known non-matches. It can be seen that the distribution of each feature for the known non-matches is quite symmetric, while the distribution for each feature for the known matches is skewed left.'}
cor_compare <- CCFs_withlands %>%
  select(ccf, rough_cor, match) %>%
  gather(key = feature, value = value, ccf, rough_cor) %>%
  mutate(match = factor(match, labels = c("Known non-matches (KNM)", "Known matches (KM)")))

custom <- list(
  theme_bw(),
  scale_linetype_manual("Bullet Land Pairs", values=3:2),
  scale_colour_manual("Bullet Land Pairs", values=c("black", "grey40"))
  )

ggplot(data = cor_compare, aes(x = value, colour=match, linetype = match)) +
  stat_ecdf() +
  facet_wrap(~feature, nrow = 2) +
  ylab("") + xlab("") +
  theme(legend.background = element_rect(colour="grey75"),
        axis.title.y=element_blank(),
        plot.margin = unit(c(0,0,0,0), unit="cm")) +
  custom +
    theme(legend.position=c(0.2, 0.85)) 
```

@@

CCFs_test_normalized <- CCFs_test %>% select(Type, ccf, rough_cor, D, matches, mismatches, cms, non_cms, sum_peaks)
CCFs_test_normalized[,-1] <- lapply(CCFs_test_normalized[,-1], function(x) { (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE) })
pcp_data <- CCFs_test_normalized %>%
  group_by(Type) %>%
  summarise_each(funs(mean))
pcp_data_tall <- pcp_data %>%
  gather(key = Feature, value = Value, 2:ncol(.))
CCFs_tall <- CCFs_test_normalized %>%
  mutate(id = 1:nrow(.)) %>%
  select(id, everything()) %>%
  gather(key = Feature, value = Value, 3:ncol(.)) %>%
  mutate(x = as.numeric(as.factor(Feature)))

truepos <- filter(CCFs_tall, Type %in% c("True Positive"))
falsepos <- filter(CCFs_tall, Type %in% c("False Positive"))

truepos_poly <- truepos %>% 
  group_by(Feature) %>% 
  summarize(min = min(Value), max=max(Value), mean=mean(Value), x=x[1])
tpps <- truepos_poly %>% gather(Type, Result, min:max)
tpps$order <-c(1:8, 9+8:1)
tpps$x <-c(1:8, 1:8)

tpps <- tpps %>% arrange(order)

p1 <- ggplot() +
  geom_polygon(aes(x = x, y=Result), data = tpps, fill="grey50", alpha=.35) +
  geom_line(aes(x = x, y = mean), data = truepos_poly, colour="grey50", size=1) +
  geom_line(data = truepos, aes(x = x, y = Value, colour = Type, group = id, linetype=Type), alpha = 0.1, size = 0.2) + #, linetype = "dashed") +
  scale_x_continuous(breaks=1:8, labels=unique(tpps$Feature)) +
  geom_line(data = falsepos, aes(x = x, y = Value, colour = Type, group = id, linetype=Type), alpha = 1, size = 0.8) +
  ylab("Normalized Value") +
  xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = c("grey10", "grey50")) +
#  scale_color_manual(values = c("#e60000", "#000099")) +
  scale_size_manual(values = c(1, 1)) +
  scale_linetype_manual(values=2:1)

# p1 <- ggplot() +
# #  geom_polygon(aes(x = x, y=Result), data = tpps, fill="#000099", alpha=.5) +
#   geom_line(data = truepos, aes(x = Feature, y = Value, colour = Type, group = id), alpha = 0.1, size = 0.2, linetype = "dashed") +
#   geom_line(data = falsepos, aes(x = Feature, y = Value, colour = Type, group = id), alpha = 1, size = 0.8, linetype = "dashed") +
#   geom_line(data = filter(pcp_data_tall, Type %in% c("True Positive")), aes(x = Feature, y = Value, colour = Type, group = Type), size = 2) +
#   ylab("Normalized Value") +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
#   scale_color_manual(values = c("#e60000", "#000099")) +
#   scale_size_manual(values = c(1, 1))

set.seed(20170222)
mysamp1 <- sample(CCFs_tall$id[CCFs_tall$Type == "True Negative"], 226)
mysamp2 <- sample(CCFs_tall$id[CCFs_tall$Type == "False Negative"], 20)

CCFs_tall <- CCFs_tall %>% mutate(x = as.numeric(as.factor(Feature)))

trueneg_sample <- CCFs_tall %>%
  filter(id %in% mysamp1) 
falseneg_sample <- CCFs_tall %>% 
  filter(id %in% mysamp2)

trueneg_poly <- trueneg_sample %>% 
  group_by(Feature) %>% 
  summarize(min = min(Value), max=max(Value), mean=mean(Value), x=x[1])
tnps <- trueneg_poly %>% gather(Type, Result, min:max)
tnps$order <-c(1:8, 9+8:1)
tnps$x <-c(1:8, 1:8)

tnps <- tnps %>% arrange(order)

p2 <- ggplot() +
  geom_polygon(aes(x = x, y=Result), data = tnps, fill="grey50", alpha=.35) +
  geom_line(aes(x = x, y = mean), data = trueneg_poly, colour="grey50", size=1) +
  geom_line(data = trueneg_sample, aes(x = x, y = Value, colour = Type, group = id, linetype=Type), alpha = 0.1, size = 0.2) + #, linetype = "dashed") +
scale_x_continuous(breaks=1:8, labels=unique(tpps$Feature)) +
  geom_line(data = falseneg_sample, aes(x = x, y = Value, colour = Type, group = id, linetype=Type), alpha = 1, size = 0.8) +
  ylab("Normalized Value") +
  xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = c("grey10", "grey50")) +
#  scale_color_manual(values = c("#e60000", "#000099")) +
  scale_size_manual(values = c(1, 1)) +
  scale_linetype_manual(values=2:1)

# p2 <- ggplot() +
#   #geom_line(data = filter(CCFs_tall, Type %in% c("True Negative", "False Negative")), aes(x = Feature, y = Value, colour = Type, group = id, alpha = Type, size = Type)) +
#   geom_line(data = trueneg_sample, aes(x = Feature, y = Value, colour = Type, group = id), alpha = 0.1, size = 0.2, linetype = "dashed") +
#   geom_line(data = falseneg_sample, aes(x = Feature, y = Value, colour = Type, group = id), alpha = 1, size = 0.4, linetype = "dashed") +
#   theme_bw() +
#   geom_line(data = filter(pcp_data_tall, Type %in% c("True Negative")), aes(x = Feature, y = Value, colour = Type, group = Type), size = 2) +
#   ylab("Normalized Value") +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
#   scale_color_manual(values = c("#e60000", "#000099"))

@@

## Produce plot
ggplot(result, aes(x = level, y = value, colour = fixed, linetype=fixed)) +
    facet_wrap(~metric, nrow = 2) +
    geom_point() +
    geom_line() +
    theme_bw() +
    scale_x_continuous(breaks = c(0, .125, .25, .375, .5, .625, .75), labels = c("100%", "87.5%", "75%", "62.5%", "50%", "37.5%", "25%")) +
    xlab("Land Proportion") +
  scale_colour_manual(values=c("black", "grey20", "grey30", "grey40")) +
  theme(legend.key.width=unit(3,"line"))
```

@@

ggplot(data = feature_exp, aes(x = level, y = value, colour = fixed, linetype=fixed)) +
    geom_point() +
    geom_line() +
    theme_bw() +
    theme(legend.position = "bottom") +
    scale_x_continuous(breaks = c(0, .25, .5, .75), labels = c("100%", "75%", "50%", "25%")) +
    xlab("Land Proportion") +
    facet_wrap(~feature, scales = "free_y") +
    scale_colour_manual(values=c("black", "grey20", "grey30", "grey40")) +
  theme(legend.key.width=unit(3,"line"))

```

@@

ggplot(data = feature_exp2, aes(x = level, y = value, colour = fixed, linetype=fixed)) +
    geom_point() +
    geom_line() +
    theme_bw() +
    theme(legend.position = "bottom") +
    scale_x_continuous(breaks = c(0, .25, .5, .75), labels = c("100%", "75%", "50%", "25%")) +
    xlab("Land Proportion") +
    facet_wrap(~feature, scales = "free_y") +
    scale_colour_manual(values=c("black", "grey20", "grey30", "grey40")) +
  theme(legend.key.width=unit(3,"line"))

```

@@

ggplot(data = result, aes(x = match, y = score, fill = match)) +
    geom_boxplot() +
    coord_flip() +
    scale_fill_brewer("Bullet Land Pairs", palette="Paired") +
    theme_bw() +
    theme(legend.position="none",
          axis.text.y = element_blank(),
          axis.title.y = element_blank()) +
  annotate("text", x=2.25, y= 0.8, label="Known matches (KM)", hjust=1, colour="grey40") +
  annotate("text", x=0.75, y= 0.05, label="Known non-matches (KNM)", hjust=0, colour="grey40")

```

@@

>>>>

p1 <- ggplot(data = final_profs, aes(x = y, y = l30, colour = profile_id)) +
    geom_line() +
    theme_bw()

p2 <- ggplot(data = final_doublesmoothed, aes(x = y, y = l30, colour = profile_id)) +
    geom_line() +
    geom_smooth(inherit.aes = FALSE, aes(x = y, y = avgl30), colour = "purple") +
    theme_bw()

p3 <- ggplot(data = final_doublesmoothed, aes(x = y, y = l50, colour = profile_id)) +
    geom_line() +
    theme_bw()

@@

```{r roughcordist, echo=FALSE, fig.cap='Empirical CDFs of the Roughness Correlation compared to the CCF for known matches and known non-matches. It can be seen that the distribution of each feature for the known non-matches is quite symmetric, while the distribution for each feature for the known matches is skewed left.'}
cor_compare <- CCFs_withlands %>%
  select(ccf, rough_cor, match) %>%
  gather(key = feature, value = value, ccf, rough_cor) %>%
  mutate(match = factor(match, labels = c("Known non-matches (KNM)", "Known matches (KM)")))

ggplot(data = cor_compare, aes(x = value, color = match)) +
  stat_ecdf() +
  facet_wrap(~feature, nrow = 2) +
  theme_bw() + ylab("") + xlab("") +
  scale_color_brewer("Bullet Land Pairs", palette="Paired") +
  theme(legend.background = element_rect(colour="grey75"),
        axis.title.y=element_blank(),
        plot.margin = unit(c(0,0,0,0), unit="cm"))
```

@@

CCFs_test_normalized <- CCFs_test %>% select(Type, ccf, rough_cor, D, matches, mismatches, cms, non_cms, sum_peaks)
CCFs_test_normalized[,-1] <- lapply(CCFs_test_normalized[,-1], function(x) { (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE) })
pcp_data <- CCFs_test_normalized %>%
  group_by(Type) %>%
  summarise_each(funs(mean))
pcp_data_tall <- pcp_data %>%
  gather(key = Feature, value = Value, 2:ncol(.))
CCFs_tall <- CCFs_test_normalized %>%
  mutate(id = 1:nrow(.)) %>%
  select(id, everything()) %>%
  gather(key = Feature, value = Value, 3:ncol(.))

p1 <- ggplot() +
  geom_line(data = filter(CCFs_tall, Type %in% c("True Positive")), aes(x = Feature, y = Value, colour = Type, group = id), alpha = 0.1, size = 0.2, linetype = "dashed") +
  geom_line(data = filter(CCFs_tall, Type %in% c("False Positive")), aes(x = Feature, y = Value, colour = Type, group = id), alpha = 1, size = 0.8, linetype = "dashed") +
  geom_line(data = filter(pcp_data_tall, Type %in% c("True Positive")), aes(x = Feature, y = Value, colour = Type, group = Type), size = 2) +
  ylab("Normalized Value") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = c("#e60000", "#000099")) +
  scale_size_manual(values = c(1, 1))

set.seed(20170222)
mysamp1 <- sample(CCFs_tall$id[CCFs_tall$Type == "True Negative"], 226)
mysamp2 <- sample(CCFs_tall$id[CCFs_tall$Type == "False Negative"], 20)
trueneg_sample <- CCFs_tall %>%
  filter(id %in% mysamp1)
falseneg_sample <- CCFs_tall %>% 
  filter(id %in% mysamp2)

p2 <- ggplot() +
  #geom_line(data = filter(CCFs_tall, Type %in% c("True Negative", "False Negative")), aes(x = Feature, y = Value, colour = Type, group = id, alpha = Type, size = Type)) +
  geom_line(data = trueneg_sample, aes(x = Feature, y = Value, colour = Type, group = id), alpha = 0.1, size = 0.2, linetype = "dashed") +
  geom_line(data = falseneg_sample, aes(x = Feature, y = Value, colour = Type, group = id), alpha = 1, size = 0.4, linetype = "dashed") +
  theme_bw() +
  geom_line(data = filter(pcp_data_tall, Type %in% c("True Negative")), aes(x = Feature, y = Value, colour = Type, group = Type), size = 2) +
  ylab("Normalized Value") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = c("#e60000", "#000099"))

@@

## Produce plot
ggplot(result, aes(x = level, y = value, colour = fixed)) +
    facet_wrap(~metric, nrow = 2) +
    geom_point() +
    geom_line() +
    theme_bw() +
    scale_x_continuous(breaks = c(0, .125, .25, .375, .5, .625, .75), labels = c("100%", "87.5%", "75%", "62.5%", "50%", "37.5%", "25%")) +
    xlab("Land Proportion")
```

@@

ggplot(data = feature_exp, aes(x = level, y = value, colour = fixed)) +
    geom_point() +
    geom_line() +
    theme_bw() +
    theme(legend.position = "bottom") +
    scale_x_continuous(breaks = c(0, .25, .5, .75), labels = c("100%", "75%", "50%", "25%")) +
    xlab("Land Proportion") +
    facet_wrap(~feature, scales = "free_y")
```

@@

ggplot(data = feature_exp2, aes(x = level, y = value, colour = fixed)) +
    geom_point() +
    geom_line() +
    theme_bw() +
    theme(legend.position = "bottom") +
    scale_x_continuous(breaks = c(0, .25, .5, .75), labels = c("100%", "75%", "50%", "25%")) +
    xlab("Land Proportion") +
    facet_wrap(~feature, scales = "free_y")
```

@@

ggplot(data = result, aes(x = match, y = score, fill = match)) +
    geom_boxplot() +
    coord_flip() +
    scale_fill_brewer("Bullet Land Pairs", palette="Paired") +
    theme_bw() +
    theme(legend.background = element_rect(colour="grey75"),
          axis.text.y = element_blank(),
          axis.title.y = element_blank())
```

@@

This research was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.
